### 2021-11-30 plos_latex_template.tex
RJ: Accepted many of Janani's suggestions for the Intro

The authors of this article recently organized useR! 2021, the annual conference of the R Project for Statistical Computing. useR! conferences are non-profit events organized by volunteers from the R community and managed by the R Foundation. The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector. For 2021, we aimed to build a high-quality virtual and explicitly global conference providing a kind, inclusive, accessible, and welcoming environment for everyone. 
In this article, we outline our most important lessons learned in 10 simple rules to host an inclusive conference. These rules apply equally to academic, industry, or mixed conferences; while inspired by a global experience, these rules also apply to regional events. These rules were implemented carefully and intentionally during a virtual conference, but are designed to serve as useful guidelines for in-person and hybrid events as we exit the pandemic.

@@

% Rocío: Main message of the paragraph: Conferences are great, but for some
Conferences, from the Latin \textit{conferre} `to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
The larger the conference, the larger the opportunities to network and learn from the cohort.
However, opportunities for participating in conferences are not equal for all. 
Although unintentionally, many academic and tech conferences have become exclusive spaces for privileged groups, \textit{i.e.}, with unearned advantages given by society to some but not all (\textit{e.g.}, white, male, from a rich country, English-native speaker, with no physical disabilities) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
% However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.

This article proposes 10 simple rules to host inclusive conferences.
They are organized in three sections: \textbf{the pillars} (embracing diversity, creating a welcoming environment, and having a diverse organizing team), \textbf{design rules} (counteracting bias in the conference program, embracing a strong online component, achieving accessibility for people with disabilities, embracing language diversity, enforcing inclusive communication strategies, and allocating financial resources to support inclusive practices), and a \textbf{process rule} to make the conference part of a long-term commitment to inclusion. 
These practices stem from the experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible.

The authors of this article include the global coordinators of the conference, members of the diversity and inclusion team, and people who contributed to other critical areas of the conference. 
Therefore, the rules result from our lessons learned from diverse perspectives and expertise (\textit{i.e.}, our different backgrounds, profiles, and roles in the conference). 
Even though these rules were designed based on our experience with a global event, they can be easily transferred or scaled to events with regional or local scopes.
These rules will greatly benefit people who are part of meeting committees that oversee the site/location selection process, or that coordinates with the local organizers of conferences.
The rules will also help local/virtual organizers who intend to put inclusion at the core of their conference right from the conception and planning phases. We have designed them for wide adoption to enable and ensure that the scientific and technical conferences post-pandemic are as accessible and inclusive as they can be.

@@

\begin{figure}[!h]
\centering
%sara: without success i'm trying to use pdfcomment package and pdftooltip to make the alt-text. any help?

\pdftooltip{
\includegraphics[width=\textwidth]{figs/rules.png}}{
%sara: suggestion for alt text:
%Diagram of how the 10 rules are organized into three groups: grounding rules (rules 1, 2, and 10), people-related rules (rules 3 and 4), and design rules (rules 5 to 9). The diagram has five rows. Each line contains rectangular boxes with the rule number and a short title for each rule. Each rule box is colored based on their group (grounding rules are grey, people-related rules are green and design rules are pink). First row is for the grounding rules 1. Multidimensional diversity and 2. Safe and inclusive environment in grey boxes. Second row is for people-related rules 3. Organizing team and 4. Unbias in green boxes. Third row is for the design rules 5. Online component, 6. Accessibility for disabilities, 7. Language inclusivity in pink boxes. The fourth row is for the design rules 8. Communication and 9. Financial support and budget in pink boxes. The last row is for the grounding rule 10. Diversity and inclusion as a process in grey boxes.
}
\caption{Diagram of the rules organized in three groups: the pillars (rules 1, 2, and 3), design rules (rules 4 to 9), and the process (rule 10).}
\label{fig:diagram}
\end{figure}

@@

% Rocío: more effort towards including the more excluded
This may translate into different actions depending on your field, region, or community, \textit{e.g.}, a gender distribution of your speakers that is representative of the general population, presence of racialized people — especially Black people — among the head organizers, speakers, and attendees, LGBTQIA+ friendly-spaces, or community participation from key geographic regions.

@@

>>>>

The authors of this article recently organized useR! 2021, the annual conference of the R Project for Statistical Computing. useR! conferences are non-profit events organized by volunteers from the R community and managed by the R Foundation. The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector. For 2021, we aimed to build a high-quality virtual and explicitly global conference creating a kind, inclusive, accessible, and welcoming environment for everyone. 
In this article, we outline our most important lessons learned in 10 simple rules to host an inclusive conference. These rules apply equally to academic, industry, or mixed conferences; while inspired by a global experience, these rules also apply to regional events. These rules were implemented carefully and intentionally during a virtual conference, but are designed to serve as useful guidelines for in-person and hybrid events even after the covid pandemic.

@@

% Rocío: Main message of the paragraph: Conferences are great, but for some
Conferences, from the Latin \textit{conferre} `to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
The larger the conference, the larger the opportunities to network and learn from the cohort.
However, opportunities for participating in conferences are not equal for all. 
Many academic and tech conferences have become exclusive spaces for privileged groups, \textit{i.e.}, with unearned advantages given by society to some but not all (\textit{e.g.}, white, male, from a rich country, English-native speaker, with no physical disabilities) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
% However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.

This article proposes 10 simple rules to host inclusive conferences.
They are organized in three sections: \textbf{the pillars}, fundamental rules with key elements to conceive the work on diversity and inclusion in any conference (embracing diversity, creating a welcoming environment, and having a diverse organizing team); \textbf{design rules}, concerning diverse elements in the design of the conference and operational tasks (counteracting bias in the conference program, embracing a strong online component, achieving accessibility for people with disabilities, embracing language diversity, enforcing inclusive communication strategies, and allocating financial resources to support inclusive practices); and a \textbf{process rule} to make the conference part of a long-term commitment to inclusion. 
These practices stem from the experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible.

The authors of this article include the global coordinators of the conference, members of the diversity and inclusion team, and people who contributed to other areas of the conference. 
Therefore, the rules result from our lessons learned from diverse perspectives and expertise (\textit{i.e.}, our different backgrounds, profiles, and roles in the conference). 
Even though these rules were designed based on our experience with a global event, they can be easily transferred or scaled to events with regional or local scopes.
These rules will greatly benefit people who are part of meeting committees that oversee the site/location selection process, or that coordinates with the local organizers of conferences.
The rules will also help local/virtual organizers who intend to put inclusion at the core of their conference right from the conception and planning phases. We have designed them for wide adoption to enable and ensure that the scientific and technical conferences during and post-pandemic are as accessible and inclusive as they can be.

@@

\begin{figure}[!h]
\centering
%sara: without success i'm trying to use pdfcomment package and pdftooltip to make the alt-text. any help?
%ast: R/ Unfortunately, support for adding accessibility metadata to LaTeX documents is limited. You will probably need to use Adobe Acrobat to add missing accessibility metadata to your PDF file.https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#latex
\pdftooltip{
\includegraphics[width=\textwidth]{figs/rules.png}}{
%sara: suggestion for alt text:
%Diagram of how the 10 rules are organized into three groups: grounding rules (rules 1, 2, and 10), people-related rules (rules 3 and 4), and design rules (rules 5 to 9). The diagram has five rows. Each line contains rectangular boxes with the rule number and a short title for each rule. Each rule box is colored based on their group (grounding rules are grey, people-related rules are green and design rules are pink). First row is for the grounding rules 1. Multidimensional diversity and 2. Safe and inclusive environment in grey boxes. Second row is for people-related rules 3. Organizing team and 4. Unbias in green boxes. Third row is for the design rules 5. Online component, 6. Accessibility for disabilities, 7. Language inclusivity in pink boxes. The fourth row is for the design rules 8. Communication and 9. Financial support and budget in pink boxes. The last row is for the grounding rule 10. Diversity and inclusion as a process in grey boxes.
}
\caption{Diagram of the rules organized in three groups: the pillars (rules 1, 2, and 3), design rules (rules 4 to 9), and the process (rule 10).}
\label{fig:diagram}
\end{figure}

@@

% Rocío: more effort towards including the more excluded
This may translate into different actions depending on your field, region, or community, \textit{e.g.}, a gender distribution of your speakers that is representative of the general population, presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees, LGBTQIA+ friendly-spaces, or community participation from key geographic regions.

@@

\begin{figure}[!h]
\centering
\pdftooltip{
\includegraphics[width=\textwidth]{figs/marmots.pdf}}{
}
\caption{Margot the marmot was useR! 2021 mascot}
\label{fig:marmots}
\end{figure}
